
---
title: "Thermal Stress"
output:
  pdf_document: default
  html_document: default
---

Exploration of traits that reduce thermal stress in Austrlia.
4 data files needed to run code:
  Traits
  Choa raster
  Occurange matrix
  Average annual VPD

```{r,warning=FALSE,message=FALSE}


rm(list = ls())

library(raster)
library(RColorBrewer)
library(data.table)
library(maptools)
#library(letsR)
#library(plantecophys)
library(car)
#library(mgcv)
library(ncf)
library(spdep)
#library(resample)
library(lme4)
library(plyr)
library(multcomp)
library(Hmisc)
library(MuMIn)

#Albers equal area projection
Albers<-"+proj=aea +lat_1=-18 +lat_2=-36 +lat_0=0 +lon_0=134 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"


resetPar <- function() {
    dev.new()
    op <- par(no.readonly = TRUE)
    dev.off()
    op
}

asNumeric <- function(x) as.numeric(as.character(x))
factorsNumeric <- function(d) modifyList(d, lapply(d[, sapply(d, is.factor)], asNumeric))


SACPrepMoranDistance<-function (lmModels, distkm, neighborStyle,coords){
  disNB<-dnearneigh(coords, 0, distkm,longlat=TRUE) #turns neighbourhood object into a weighted list
  listWeight<-nb2listw(disNB, glist=NULL, style=neighborStyle, zero.policy=FALSE)
  #moran.test(residuals(lmModels), listw=listWeight)
  SARlm <- errorsarlm(lmModels, listw=listWeight)
  #summary(SARlm, Nagelkerke=TRUE) #Gives a summary of the SAR error model

  coeffcnt<-as.numeric(SARlm$coefficients[2])
  SE<- as.numeric(SARlm$rest.se[2])
  lmAIC<-SARlm$AIC_lm.model # this is AIC of lm
  sarAIC<-AIC(SARlm)
  NagelkerkePseudoRsquared<-summary(SARlm, Nagelkerke=TRUE)[[47]]
  #get values  of Moran’s I over the first twenty distance groupings
  correlo <- correlog(coords[,1], coords[,2], residuals(SARlm),
na.rm=T, increment=1, resamp=0)
  MoransI20distGroup <-sum(correlo$correlation[1:20])
  outdat<-t(c(distkm,neighborStyle,lmAIC,sarAIC,MoransI20distGroup,coeffcnt,SE,NagelkerkePseudoRsquared))
  colnames(outdat)<-c("SACdistance","neighborStyle","lmAIC","sarAIC","MoransI20distGroup","regression coefficients","SE","NagelkerkePseudo-R-squared")
 return(outdat)
}

#colours for graphs
rgb.palette <-
        colorRampPalette(c("darkgreen",
                  "orange",
                  "darkblue"),
                  space = "rgb")

#########################
#load needed data
#########################
#traits
traits<-read.csv('/Users/daisy/Google Drive/PhD/ThermalStress/tables/Table_S1_20161128.csv')
          #occurence records
          # Occ<-read.csv("~/Google Drive/PhD/ThermalStress/data/PassarineObservationsGriddedFinal2016-06-21.csv")
          # Occ<-unique(Occ)
#read in observations
Occ <- fread("~/Google Drive/PhD/ThermalStress/data/PassarineObservationsFinal2016-06-21.csv",
                        data.table=FALSE)
#change XY to albers equal area
Locs_sp <- SpatialPoints(cbind(Occ$lon, Occ$lat), proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84"))
Locs_sp <- spTransform(Locs_sp, CRS=CRS(paste(Albers)))
Alberslocs<-as.data.frame(Locs_sp)
colnames(Alberslocs) <- c("Lon_Albers","Lat_Albers")
Occ<-cbind(Occ,Alberslocs)

# species <- unique(occ$Scientific.Name)




#completeness index
#ChaoR<-raster("/Users/daisy/Google Drive/PhD/ThermalStress/data/CmpltDomeCupGridCells100km2017-02-14.asc",crs=Albers)
ChaoR<-raster("/Users/daisy/Google Drive/PhD/ThermalStress/data/ChoaIndex100km10SpeciesMinimum2017-02-15.asc",crs=Albers)
#ChaoR<-trim(ChaoR)
Chao<-rasterToPoints(ChaoR)
# 100km mask of Austrlaia where each gridcell has unique ID
aus<-ChaoR
aus[]<-1:1480 #Give each grid cell unique value
aus<-raster::mask(aus,ChaoR)#keep only cells interested in
#remove Occ that do not occur in gridcells we want to asses, CI < 0.7
Occ<-unique(Occ[,c('Scientific.Name','Lon_Albers',"Lat_Albers")])
Occ$ID<-raster::extract(aus,cbind(Occ$Lon_Albers,Occ$Lat_Albers))#Grid cell ID for each presence record
Occ<-subset(Occ,ID!= "NA")
GridOcc<-unique(Occ[c("Scientific.Name","ID")])
#summary of gidcell occupancey
occTable<-as.data.frame(table(GridOcc$ID))
min(occTable$Freq)#min number of species per gridcell
max(occTable$Freq)#max number of species per gridcell
mean(occTable$Freq)#mean number of species per gridcell
median(occTable$Freq)#mean number of species per gridcell
spTable<-as.data.frame(table(GridOcc$Scientific.Name))
min(spTable$Freq)#min number of grid cells occupied by a species
max(spTable$Freq)#max number of gridcells occupied by a species
mean(spTable$Freq)
median(spTable$Freq)

##################################

#Figure of Species Richness

#################################

RichAllSPecies<-raster::subs(aus,occTable)
rgb.palette <-
  colorRampPalette(c("darkgreen",
                     "orange",
                     "darkblue"),
                   space = "rgb")

ausOutline <- readShapeSpatial("/Users/daisy/Google Drive/PhD/Data/Spatial/OutlineOfAustralia/Australia_original_nostates.shp", proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84"))
ausOutline <-spTransform(ausOutline, CRS=CRS(paste(Albers)))


pdf(file = paste0("/Users/daisy/Google Drive/PhD/ThermalStress/figures/manuscript/SpeciesRichenss",as.Date(Sys.time()),".pdf"),
    width = 3, height = 3)
par(mar=c(0,0,0,0))
plot(ausOutline)
plot(RichAllSPecies,col=rgb.palette(9),legend=FALSE, axes=FALSE,box=FALSE,add=TRUE)
plot(ausOutline, bg="transparent", add=TRUE,border="black")
plot(RichAllSPecies,
     legend.only=T,
     col=rgb.palette(9),
     smallplot=c(0.08,0.12, 0.1,0.23),
     axis.args=list(cex.axis=.8,
                    at=c(40,80,120,160),
                    labels=c(40,80,120,160)))
 dev.off()


```
Load in VPD and LAI and make figure

```{r, echo=FALSE}
#mask of Australia based on BOM data
ausGrid<-raster::raster("/Users/daisy/Google Drive/PhD/ThermalStress/data/spatial/aus100kmMask.asc",
                    crs = Albers)#mask of Australia
#VPD
VPDann<-raster("/Users/daisy/Google Drive/PhD/ThermalStress/data/spatial/VPD100km2017-02-13.asc")
VPDann<-raster::trim(raster::mask(VPDann,ausGrid))
#LAI
LAIann<-raster("/Users/daisy/Google Drive/PhD/ThermalStress/data/spatial/LAI100km2017-02-13.asc")
LAIann<-raster::trim(raster::mask(LAIann,ausGrid))

##############################

#Figure - VPD and LAI

#############################

pdf(file = paste0("/Users/daisy/Google Drive/PhD/ThermalStress/figures/manuscript/VPDandLAI",as.Date(Sys.time()),".pdf"),
    width = 7.48, height = 3)

rgb.palette <-
  colorRampPalette(c("darkgreen",
                     "orange",
                     "darkblue"),
                   space = "rgb")

par(mfrow=c(1,2),
    mar=c(0,1,0,0))
plot(ausOutline)
plot(VPDann,col=rgb.palette(9),legend=FALSE, axes=FALSE,box=FALSE,ext=extent(trim(aus)),add=TRUE)
plot(ausOutline, bg="transparent", add=TRUE,border="black")
plot(VPDann,
     legend.only=T,
     col=rgb.palette(9),
     smallplot=c(0.08,0.12, 0.04,0.23),
     axis.args=list(cex.axis=.8,
                    at=c(1:4),
                    labels=c(1:4)),
    legend.args=list(text='VPD (kPa)', side=3, font=2, line=1, cex=0.8))
legend("topleft" , "(a)", bty="n",text.font=c(font=2),cex=1) 

#LAI

plot(ausOutline)
plot(LAIann,col=rgb.palette(9),legend=FALSE, axes=FALSE,box=FALSE,ext=extent(trim(aus)),add=TRUE)
plot(ausOutline, bg="transparent", add=TRUE,border="black")
plot(LAIann,
     legend.only=T,
     col=rgb.palette(9),
     smallplot=c(0.08,0.12, 0.04,0.23),
     axis.args=list(cex.axis=.8,
                    at=seq(0.5,2.5,1),
                    labels=seq(0.5,2.5,1)),
    legend.args=list(text='LAI', side=3, font=2, line=1, cex=0.8))
legend("topleft" , "(b)", bty="n",text.font=c(font=2),cex=1) 


dev.off()




```

Summary Stats and explore data

```{r, echo=FALSE}


#####################
#get summary values for elongation
####################

length(na.omit(traits$Elongation))
min(na.omit(traits$Elongation))
max(na.omit(traits$Elongation))
mean(na.omit(traits$Elongation))
sd(na.omit(traits$Elongation))
hist(na.omit(traits$Elongation))
sd(na.omit(traits$Elongation))/sqrt(length(na.omit(traits$Elongation))) #standard error
## Check if normally distributed
qqPlot(na.omit(traits$Elongation))

```

Check if within grid cells it is normally distributed

```{r, echo=FALSE}

locs<-unique(Occ$ID)
message(paste0("Number of 100km grid cells being used = ",length(locs)))
message(paste0("Number of Passarine species in study = ",length(unique(Occ$Scientific.Name))))

IDs<-sample(locs, 40)
#loop through cells and get average trait values
for(j in 1:length(IDs)){ 
  sp<-subset(GridOcc,ID==IDs[j])$Scientific.Name
  #get species level data
  tr<-traits[traits$Species %in% sp,]#pull out the traits for the species of interest
  tr<-na.omit(tr[,"Elongation"])
  qqPlot(tr, main =subset(Occ,ID==IDs[j])[1,c("Lon_Albers","Lat_Albers")] )#data is normal if it falls within the confident interval
}
# the data is normally distributed, but the species that have the roundedest shape sometimes fall outside of the qqPlots confidence intervals

```

Analysis 1 - similarity of egg elongation for species with different nest-types

```{r, echo=FALSE}


#elongation by nest type
nest.type<-ddply(traits,~Nest.type,summarise,mean=mean(na.omit(Elongation)),
                 sd=sd(na.omit(Elongation)),n=length(na.omit(Elongation)))
plot(Elongation ~ Nest.type, data=traits)
#Cavity significantly differnt?
mod1<-lm(Elongation ~ Nest.type,data=traits)
qqPlot(residuals(mod1),main=("mod1"))
summary(mod1)
Anova(mod1, test.statistic="F")
#pair wise comparisons
TukeyNest<- glht(mod1, linfct=mcp(Nest.type="Tukey"))
summary(TukeyNest)

#mixed effect model returns same results

mod1mx<-lmer(Elongation ~ Nest.type + (1|Family) ,data=traits)
qqPlot(residuals(mod1mx),main=("mixed effect"))
summary(mod1mx)
Anova(mod1mx, test.statistic="F")
#pair wise comparisons
TukeyNest<- glht(mod1mx, linfct=mcp(Nest.type="Tukey"))
summary(TukeyNest)
r.squaredGLMM(mod1mx)


```

Make new dataframe wtih the average values for each gridcell 
-elongation cupped nests
-elongation domed nests
-proportion domed nests
```{r, echo=FALSE}

#loop through cells and get average trait values
celltraitPass<-list()
for(lc in 1:length(locs)){ 
  sp<-subset(GridOcc,ID==locs[lc])$Scientific.Name
  if(length(sp)<=10){
    stop("Not enough species")
  }
  tr<-traits[traits$Species %in% sp,]#pull out the traits for the species of interest
  eggElong<-mean(na.omit(tr$Elongation))
  eggElongDome<-mean(na.omit(subset(tr,Nest.type=="Dome",select=Elongation)[,1]))#mean elongation for domed
  eggElongCup<-mean(na.omit(subset(tr,Nest.type=="Cup",select=Elongation)[,1]))#mean elongation for cupped
  CupRichness<-nrow(subset(tr,Nest.type=="Cup"))
  DomeRichness<-nrow(subset(tr,Nest.type=="Dome"))
  Richness<-CupRichness+DomeRichness
  propDome<-DomeRichness/(DomeRichness+CupRichness)
    celltraitPass[[lc]]<-c(locs[lc],
                      eggElong,
                      eggElongDome,
                      eggElongCup,
                      CupRichness,
                      DomeRichness,
                      Richness,
                      propDome)
}
celltraitPass<-as.data.frame(do.call("rbind",celltraitPass))
colnames(celltraitPass)<-c("ID",
                      "eggElong",
                      "eggElongDome",
                      "eggElongCup",
                      "CupRichness",
                      "DomeRichness",
                      "Richness",
                      "propDome")


```
Exploratory figure of  Richness for diffenent nest types 

```{r, echo=FALSE}

RichAllSpecies<-subs(aus,
             data.frame(celltraitPass[,c("ID","Richness")]))
RichDomed<-subs(aus,
             data.frame(celltraitPass[,c("ID","DomeRichness")]))
RichCup<-subs(aus,
             data.frame(celltraitPass[,c("ID","CupRichness")]))
domedNests<-subs(aus,
             data.frame(celltraitPass[,c("ID","propDome")]))

elongCup<-subs(aus,
              data.frame(celltraitPass[,c("ID","eggElongCup")]))
cellStats(elongCup,'mean')
 
elongDomed<-subs(aus,
              data.frame(celltraitPass[,c("ID","eggElongDome")]))

par(mfrow=c(1,3))
breakpoints <- seq(10,150,10)

colors <- (rgb.palette(14))
plot(RichAllSpecies,breaks=breakpoints,col=colors, main = "All Species")
plot(RichDomed,breaks=breakpoints,col=colors, main = "Domed")
plot(RichCup,breaks=breakpoints,col=colors, main = "Cup")

```
compare extreme gridcells

```{r, echo=FALSE}
#find extreme VPD
ExVPD<-VPDann
#arid
ExVPD[ExVPD <= 4.4] <- NA
exLAI<- (ExVPD/ExVPD)*LAIann
plot(exLAI)
exLAI<-raster::mask(exLAI,aus)
exID<-(exLAI/exLAI)*aus 
aridID<-rasterToPoints(exID)[,"layer"]
#wet
minVPD<-VPDann
minVPD[minVPD >= 1.6] <- NA
plot(minVPD)
MinLAI<- (minVPD/minVPD)*LAIann
plot(MinLAI)
MinLAI[MinLAI <= 2] <- NA
plot(MinLAI)
minID<-(MinLAI/MinLAI)*aus 
minID[minID > 600] <- NA
MinLAI<-MinLAI*(minID/minID)
minVPD<-minVPD*(minID/minID)
plot(minVPD)
wetID<-(minVPD/minVPD)*aus 
wetID<-rasterToPoints(wetID)[,"layer"]

#get species data
dry<-unique(GridOcc[GridOcc$ID %in% aridID,]$Scientific.Name)
wet<-unique(GridOcc[GridOcc$ID %in% wetID,]$Scientific.Name)
trDry<-traits[traits$Species %in% dry,]
trDry$aridity<-"dry"
trWet<-traits[traits$Species %in% wet,]
trWet$aridity<-"wet"
dryWetDat<-rbind(trDry,trWet)

dryWetDat$aridity<-as.factor(dryWetDat$aridity)
dryWetDat<-subset(dryWetDat,Nest.type != "Cavity")
#compare using mixed effect model
summary(aridLM<-lmer(Elongation ~ aridity + Nest.type + (1|Family),data=dryWetDat))
Anova(aridLM, test.statistic="F")
#find differencs between regions
TukeyRegion1<- glht(aridLM, linfct=mcp(aridity="Tukey"))
TukeyRegion2<- glht(aridLM, linfct=mcp(Nest.type="Tukey"))
summary(TukeyRegion1)
summary(TukeyRegion2)

summary(aridLM<-lm(Elongation ~ aridity,data=dryWetDat))
Anova(aridLM, test.statistic="F")



```

Prep data for SAR models 


```{r}


propDOMEp<-rasterToPoints(domedNests)
elongCUPp<-rasterToPoints(elongCup)
elongDOMEp<-rasterToPoints(elongDomed)
VPDp<-rasterToPoints(VPDann)
LAIp<-rasterToPoints(LAIann)
colnames(VPDp)<-c("x","y","VPD")
colnames(LAIp)<-c("x","y","LAI")
dat<-merge.data.frame(propDOMEp,elongCUPp,by=c("x","y"))
dat<-merge.data.frame(dat,elongDOMEp,by=c("x","y"))
dat<-merge.data.frame(dat,VPDp,by=c("x","y"))
dat<-merge.data.frame(dat,LAIp,by=c("x","y"))

```

Analysis 2 and 3 - 
(2) variation in egg elongation in relation to VPD and LAI 
(3) variation in the proportion of species with enclosed nests in relation to VPD and LAI. 


```{r}

# linear models
summary(lmElCup <- lm(eggElongCup ~ VPD * LAI, data=dat))
summary(lmElDome <- lm(eggElongDome ~ VPD * LAI, data=dat))
summary(lmPropDome <- lm(propDome ~ VPD * LAI, data=dat))



# test of spatial autocorrelation
GridCoords<-as.matrix(cbind(dat$x,dat$y))
nlist<-dnearneigh(GridCoords,d1=0,d2=200000) #distances are in meters, 
listWeight<-nb2listw(nlist, style="W")
moran.test(residuals(lmElCup), listw=listWeight)
moran.test(residuals(lmElDome), listw=listWeight)
moran.test(residuals(lmPropDome), listw=listWeight)

#The resulting p values from moran.test support the conclusion that the errors in models are spatially autocorrelated, run a spatial error model - simultaneous autoregressive model (SAR). Run SAR model with weighted lists

#error -  200 km, W
elongNB<-dnearneigh(GridCoords, 0, 200000,longlat=FALSE) #turns neighbourhood object into a weighted list
elonglistWeight<-nb2listw(elongNB, glist=NULL, style="W", zero.policy=FALSE)
#SAR models
summary(SARlmElCup <- errorsarlm(lmElCup, listw=elonglistWeight),
        Nagelkerke=T) #sig
summary(SARlmElDome <- errorsarlm(lmElDome, listw=elonglistWeight),
        Nagelkerke=T)#sig
summary(SARlmPropDome <- errorsarlm(lmPropDome, listw=elonglistWeight),
        Nagelkerke=T)#sig





######################

#Figure Elongation and propotion domed

######################

# #get range of VPD values under differnt low, averag, and high LAI
QLAI<-quantile(dat$LAI, c(.05,.95))
rangeVPDMinLAI<-c(min(subset(dat,LAI<=QLAI[1],select=VPD)),max(dat$VPD))

midrangedata<-subset(dat,LAI>QLAI[1] & LAI<QLAI[2],select=VPD)
rangeVPDMeanLAI<-c(min(midrangedata),max(midrangedata))

rangeVPDMaxLAI<-c(min(dat$VPD),max(subset(dat,LAI>=QLAI[2],select=VPD)))


#predict response to VPD under differnt LAI
predLAILow <- data.frame(LAI = QLAI[1],
                        VPD = seq(min(subset(dat,LAI<=QLAI[1],select=VPD)),
                                  max(dat$VPD),length=101))
predLAImean <- data.frame(LAI = mean(dat$LAI),
                        VPD = seq(rangeVPDMeanLAI[1],rangeVPDMeanLAI[2], length=101))
predLAIHigh <- data.frame(LAI = QLAI[2],
                        VPD = seq(rangeVPDMaxLAI[1],rangeVPDMaxLAI[2], length=101))

predLAILow$Domepred <- predict(SARlmElDome, newdata=predLAILow)
predLAILow$CuppredM <- predict(SARlmElCup, newdata=predLAILow)
predLAILow$PropDome <- predict(SARlmPropDome, newdata=predLAILow)

predLAImean$Domepred <- predict(SARlmElDome, newdata=predLAImean)
predLAImean$CuppredM <- predict(SARlmElCup, newdata=predLAImean)
predLAImean$PropDome <- predict(SARlmPropDome, newdata=predLAImean)

predLAIHigh$Domepred <- predict(SARlmElDome, newdata=predLAIHigh)
predLAIHigh$CuppredM <- predict(SARlmElCup, newdata=predLAIHigh)
predLAIHigh$PropDome <- predict(SARlmPropDome, newdata=predLAIHigh)

#function to make abline with range (from: Remko Duursma)
abline_range <- function(a=NULL,b=NULL,reg=NULL,from=NULL,to=NULL,...){
  # Borrowed from abline
  if (!is.null(reg)) a <- reg
  if (!is.null(a) && is.list(a)) {
    temp <- as.vector(coefficients(a))
    from <- min(a$model[,2], na.rm=TRUE)
    to <- max(a$model[,2], na.rm=TRUE)
    if (length(temp) == 1) {
      a <- 0
      b <- temp
    }
    else {
      a <- temp[1]
      b <- temp[2]
    }
  }
  segments(x0=from,x1=to,
           y0=a+from*b,y1=a+to*b,...)
}

{
pdf(file = paste0("/Users/daisy/Google Drive/PhD/ThermalStress/figures/manuscript/ElongationProportionMapsPlots",as.Date(Sys.time()),".pdf"),
    width = 7.48, height = 10)

par(mfrow = c(3,2),          
    mar = c(2,0,2,1) + 0.1)

rgb.palette <-
  colorRampPalette(c("darkgreen",
                     "orange",
                     "darkblue"),
                   space = "rgb")
breakpoints <- seq(1.33,1.42,.01)


#(a) - Elongation for domed nests
plot(elongDomed,col=rgb.palette(9),breaks=breakpoints,axes=F,legend=F,box=F)
legend("topleft", "(a)", bty="n",text.font=c(font=2),cex=1.5)
plot(ausOutline, bg="transparent", add=TRUE,border="black")
plot(elongDomed,
     legend.only=T,
     col=rgb.palette(9),
     breaks=breakpoints,
     smallplot=c(0.08,0.12, 0.04,0.23),
     axis.args=list(cex.axis=1.2,
                    at=c(1.348,1.38,1.412),
                    labels=seq(1.34,1.42, 0.04)))

par(mar =  c(4, 4, 1, 1)) 
with(dat, plot(VPD, eggElongDome,
               pch=20,
               col="grey",
               cex=1.2,
               xlab="VPD (kPa)", 
               ylab = "Elongation - domed nests",
               frame=F,
               axis.args=list(cex.axis=1.2)))
box(which = "plot", bty = "l")
with( predLAIHigh, lines(VPD, Domepred, lwd=2, col="darkgreen"))
with(predLAILow, lines(VPD, Domepred, lwd=2, col="darkblue"))
with(predLAImean, lines(VPD, Domepred, lwd=2, col="orange"))

ELDomR2<-round(summary(SARlmElDome,Nagelkerke=T)$ NK,2) 
myr2 = bquote(italic(R)^2 == .(ELDomR2))
legend("bottomleft", legend  = myr2, bty="n",cex=1.2) 

#(b) - Elongation for cup nests
par(mar = c(2,0,2,1) + 0.1)
plot(elongCup,col=rgb.palette(9),breaks=breakpoints,axes=F,legend=F,box=F)
plot(ausOutline, bg="transparent", add=TRUE,border="black")
legend("topleft" , "(b)", bty="n",text.font=c(font=2),cex=1.5)
plot(elongCup,
    legend.only=T,
    breaks=breakpoints,
     col=rgb.palette(9),
     smallplot=c(0.08,0.12, 0.04,0.23),
     axis.args=list(cex.axis=1.2,
                    at=c(1.348,1.38,1.412),
                    labels=seq(1.34,1.42, 0.04)))


par(mar =  c(4, 4, 1, 1)) 
with(dat, plot(VPD, eggElongCup,
               pch=20,
               col="grey",
               cex=1,
               xlab="VPD (kPa)", 
               ylab = "Elongation - cupped nests",
               frame=F))
box(which = "plot", bty = "l")
with( predLAIHigh, lines(VPD, CuppredM, lwd=2, col="darkgreen"))
with( predLAImean, lines(VPD, CuppredM, lwd=2, col="orange"))
with(predLAILow, lines(VPD, CuppredM, lwd=2, col="darkblue"))

ELCupR2<-round(summary(SARlmElCup,Nagelkerke=T)$ NK,2) 
cupr2 = bquote(italic(R)^2 == .(ELCupR2))
legend("bottomleft", legend  = cupr2, bty="n") 

##Proportion covered

par(mar = c(2,0,2,1) + 0.1)
plot(domedNests,col=rgb.palette(9),axes=F,legend=F,box=F)
plot(ausOutline, bg="transparent", add=TRUE,border="black")
legend("topleft" , "(c)", bty="n",text.font=c(font=2),cex=1.5) 
plot(domedNests,
     legend.only=T,
     col=rgb.palette(9),
     smallplot=c(0.08,0.12, 0.04,0.23),
     axis.args=list(cex.axis=1,
                    at=c(.2,.3, .4),
                    labels=c(.2,.3, .4)))

par(mar =  c(4, 4, 1, 1)) 
with(dat,plot(propDome ~ VPD,
              pch=20,
              col="grey",
              cex=1,
              xlab="VPD (kPa)", 
              ylab = "Domed nest (proportion)",
              frame=F))
box(which = "plot", bty = "l")
with(predLAIHigh, lines(VPD, PropDome, lwd=2, col="darkgreen"))
with(predLAILow, lines(VPD, PropDome, lwd=2, col="darkblue"))
with( predLAImean, lines(VPD, PropDome, lwd=2, col="orange"))


ELDomR2<-round(summary(SARlmElDome,Nagelkerke=T)$ NK,2) 
PDomer2 = bquote(italic(R)^2 == .(ELDomR2) )
legend("bottomleft", legend  = PDomer2, bty="n") 

dev.off()  

}





```

```{r,echo=FALSE}


########################################

#CI scores
#function to make map CI Score using bootstrap methodo

# #######################################
# # 
# #make dataframe of unique grid cells and species richness
#  zdat<-as.data.frame(table(Occ$ID))
# 
# #species richenss map
# richness<-subs(aus, data.frame(zdat[,c("Var1","Freq")]))
# plot(richness,main = "Species richness")
# 
#  
# CIBOOT<-function (GridCell,GriddedOccurance,traitData,traitOfInterst){ 
#     sp<-as.character(subset(GriddedOccurance,ID==GridCell)[,"Species"])
#     tr<-traitData[traitData$Species %in% sp,]#pull out the traits for the species of interest
#     obs_data<-na.omit(tr[,paste(traitOfInterst)])
#     bootdat<-bootstrap(obs_data, mean(obs_data), R = 10000)
#     obs_se <- bootdat$stats$SE
#     CI<-obs_se*2 #95% CI
#     spCount<-length(sp)
#     spWithknownValue<-length(obs_data)
#     dfOUT<-cbind(GridCell,CI,spCount,spWithknownValue)
#     return(dfOUT)
# }
# 
# 
# #loop through cells and get average trait values
# CIscores<-list()
# propCIscores<-list()
# for(ID in 1:length(unique(Occ$ID))){ 
# CIscores[[ID]]<- CIBOOT(GridCell=unique(Occ$ID)[ID],
#                  GriddedOccurance=Occ,
#                  traitData=traits,
#                  traitOfInterst="Elongation")
#   
# 
# propCIscores[[ID]]<- CIBOOT(GridCell=unique(Occ$ID)[ID],
#                  GriddedOccurance=Occ,
#                  traitData=traits,
#                  traitOfInterst="roof")
# 
# message(ID)
# }
# 
# elongCIscores<-as.data.frame(do.call("rbind",CIscores))
# elongCIsore<-subs(aus, data.frame(elongCIscores[,c("GridCell","CI")]))
# plot(elongCIsore,main = "CI")
# 
#   
# PropCIscores<-as.data.frame(do.call("rbind",propCIscores))
# PropCIsoreR<-subs(aus, data.frame(na.omit(PropCIscores[,c("GridCell","CI")])))
# plot(PropCIsoreR,main = "CI")
# 
# 


################



```

# test to see if elongation changes within a species due to differences in VPD
#
# **Removed all values from QVM - the were incorrectly entered into the .csv file
#
# mixed effect model with VPD as fixed effect and nest ID as random effects
#
# "Linear mixed-effects models were used to assess how within species egg elongation changed in response to VPD. ANOVA was calculated using an Analysis of Deviance Table (Type II Wald F tests with Kenward-Roger approximation for degrees-of-freedom). In our models for each species  VPD was a fixed effect and nest ID was a random effect to account for differences in egg shape within a single nest."
#

```{r,echo=FALSE}


#species of interest
species<-c("Cracticus tibicen","Petrochelidon ariel","Zosterops lateralis",
           "Anthus australis","Rhipidura leucophrys","Grallina cyanoleuca",
           "Petrochelidon nigricans","Taeniopygia bichenovii","Pachycephala rufiventris",
           "Rhipidura albiscapa","Pomatostomus temporalis","Melanodryas cucullata",
           "Manorina melanocephala","Anthus novaeseelandiae")

#rename
spDat<-read.csv("/Users/daisy/Google Drive/PhD/ThermalStress/data/151124_otherSpecies_egg_collection_details_rearranged.csv")
  
nms<-c(names(spDat)[1:12],paste(rep(c("egg_area","egg_length","egg_width"),11),rep(1:11,each=3),sep="."),names(spDat)[46])
colnames(spDat)<-nms

#change to wide format
longDat<-list()
for (l in 1:11){
  df1<-spDat[,c(names(spDat)[1:12],paste(c("egg_area","egg_length","egg_width"),rep(l,3),sep="."),names(spDat)[46])]
  colnames(df1)<-c(names(spDat)[1:12],"egg_area","egg_length","egg_width",names(spDat)[46])
  longDat[[l]]<-df1
}
longDat<-as.data.frame(do.call(rbind,longDat))
longDat<-subset(longDat,!is.na(longDat$egg_area))#remove NA values
longDat<-subset(longDat,!is.na(longDat$egg_length))#remove NA values

#Extract VPD at the locations
longDat$VPD<-extract(VPDann,cbind(longDat$decimalLongitude,longDat$decimalLatitude))
longDat$LAI<-extract(LAI,cbind(longDat$decimalLongitude,longDat$decimalLatitude))

#for cells falling just off the coast use bilinear interpolation to get VPD values
longDatNA<-subset(longDat,is.na(longDat$VPD))
longDat<-subset(longDat,!is.na(longDat$VPD))
longDatNA$VPD<-extract(VPDann,cbind(longDatNA$decimalLongitude,longDatNA$decimalLatitude),method='bilinear')
longDatNA<-subset(longDatNA,!is.na(longDatNA$VPD))
longDat<-rbind(longDat,longDatNA)
longDat$elongation<-with(longDat,egg_length/egg_width)


#loop through each species and find out if egg elongation changes with VPD,
#we expect some species will have decreased elongation with increasing VPD
#some nests have multiple eggs so treat nest ID as random effect

modelDat<-list()
for (sp in 1:length(species)){
  df2<-subset(longDat, Species == species[sp])
  observationCount<-nrow(df2)
  uniqueNests<-length(unique(df2$id))
  #mixed effect model to control for multiple eggs in a nest
  mxElongIntra<-lmer(elongation ~ VPD * LAI + (1|id),data=df2,REML=TRUE)#add nest ID
  qqPlot(residuals(mxElongIntra),main=("mod1"))
  summary(mxElongIntra)
  Anova(mxElongIntra, test.statistic="F")
# # #check on spatial autocorrelation
#        C_lmEl <- correlog(df2$decimalLongitude, df2$decimalLatitude, residuals(mod1),
#                           na.rm=T, increment=100, resamp=0,latlon=TRUE)
#       
#        plot(0,type="n",col="black",ylab="Moran's I",xlab="lag distance",xlim=c(0,4600),ylim=c(-1,1),main = "Elongation")
       #abline(h=0,lty="dotted")
       #lines(C_lmEl$correlation~C_lmEl$mean.of.class,col="red",lwd=2)
       plot(VPDann,col="grey70",legend=FALSE,main = species[sp])
       points(df2$decimalLongitude, 
              df2$decimalLatitude, 
              col=c("blue","red")[sign(resid(mxElongIntra))/2+1.5],
              pch=19,
              cex=abs(resid(mxElongIntra))/max(resid(mxElongIntra))*2, 
              add=TRUE)
  text(120,-10,paste(sp))
  p<-round(Anova(mxElongIntra)$"Pr(>Chisq)",digits = 3)#get p values from lmer
    #text(130,-40,p)
  modAnova<-Anova(mxElongIntra, test.statistic="F")
  f<-modAnova$F
  df<-modAnova$Df
  dfRes<-modAnova$"Df.res"

  modelDat[[sp]]<-cbind(species[sp],observationCount,uniqueNests,p,f,df,dfRes)
}

do.call("rbind",modelDat)

# If I wanted to report results for sp <- 14

#F(1, 4.23) = 0.5495518, p = 0.37

# 5  F(1, 10.7) = 5.7, p < 0.05
# 10 F(1, 3.3) = 5.2, p < 0.05
# 13 F(1, 34.5) = 3.4, p < 0.05
# 



